outlier_exploration

### 2025-04-26 EDA/multicollinearity_and_outlier.qmd
Ethan: Removed aliased variable (BMI), obtained VIF values to observe multicolinearity, used step-wise AIC variable selection which removed quality_of_sleep variable, obtained VIF values of new model and were improved dramatically.

@@

>>>>

```{r}

@@

```{r}
library(car)

# Calculate VIFs on the cleaned model
vif_values <- vif(full_model_clean)

# Running VIF gave error (aliased variables are present in the model), so find out which are aliased 
alias(full_model_clean) # BMI->Overweight is perfectly aliased with other variables in the model

# Refit the model without BMI
full_model_clean2 <- lm(sleep_duration ~ gender + age + occupation + quality_of_sleep +
                        physical_activity_level + stress_level + heart_rate +
                        daily_steps + sleep_disorder + MAP, data = sleep_clean)

# Calculate VIFs on the cleaned model, with BMI removed 
vif_values <- vif(full_model_clean2)
vif_values     # Significant multicolinearity present, variable selection is next. Start with step-wise AIC variable selection

### Variable Selection

```{r}
# Step-wise AIC Variable Selection

library(MASS)

step_model <- stepAIC(full_model_clean2, direction = "both", trace = FALSE)

# Check final selected model
summary(step_model)

# Observing which variables were removed from full model 
formula(full_model_clean2) 
formula(step_model)      # Variable removed: quality_of_sleep

# Check VIF of the step-wise model
vif(step_model)      # VIF values have decreased significantly after step-wise variable selection
